### 2021-07-13 chapters/7_discussion.tex
number of clusters

In both variants we observe a reduction in average accuracy when reducing the amount of clusters.

@@

\section{RQ 2: Training a machine learning model and clustering mutant characteristics}
The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance) of the experiment, is above the statistical threshold.
The biggest deviation measured is 22.39\%.
This is significantly higher than the biggest deviation measured in the first experiment.
56\% of the results of both variants had a deviation in mutation score bigger than 5\%.
\begin{finding}
Our hypothesis does not hold for the clustering the mutants with the machine learning model trained in our experiment.
\end{finding}
We observe that the number of clusters does not have an obvious relation to the total amount of mutants. 
This is expected as the subtractive clustering does not look at the total amount of data points(mutants) but to the contents of the data points.
The number of clusters compared to the cluster amounts of the hierarchical clustering is significantly lower.
While we do not know the exact clusters centroids the subtractive model calculated, a low cluster amount indicates close relationships between certain mutants.
\begin{finding}
The mutants represented as characteristics in our data set per project have, according to the subtractive clustering model, close relationships to certain other mutants in the same data set.
\end{finding}
We observe that the minimum lowest minimum cluster accuracy measured over all variants and projects is 50.54\% with and average of 62\%. 
This is higher than the minimum of the hierarchical clustering.
This means that in all project there was not one cluster that was evenly divided.
While the minimum accuracy was higher, the maximum accuracy was lower compared to the maximum accuracy of the hierarchical clustering results.
The highest accuracy measured over all variants and projects is 97.80\% with an average of 83.79\%
This means that in all project there was not one cluster that consisted of only survived or killed mutants.
\begin{finding}
The overall accuracy of the clusters generated by the machine learning model in both variants is lower compared to that of the hierarchical clustering.
\end{finding}
When we compare the accuracy per project between the machine learning model and hierarchical clustering we can observe that the lower accuracy of the machine learning model clusters have a higher impact on the deviation in mutation score.
This is caused by the relatively low amount of clusters.
One cluster contains more mutants which results in a bigger weight per cluster.
\begin{finding}
The inaccuracy of clusters has a bigger impact on mutation score when the cluster amount is lower.
\end{finding}

>>>>

In both variants we observe a reduction in average accuracy when reducing the number of clusters.

@@

\section{RQ 2: Training a machine learning model and clustering mutant characteristics}
The maximum deviation from the original mutation score, for all variants(with and without Levenshtein distance) of the experiment, is above the statistical threshold.
The biggest deviation measured is 22.39\%.
This is significantly higher than the biggest deviation measured in the first experiment.
56\% of the results of both variants had a deviation in mutation score bigger than 5\%.
\begin{finding}
Our hypothesis does not hold for the clustering the mutants with the machine learning model trained in our experiment.
\end{finding}
We observe that the number of clusters does not have an obvious relation to the total amount of mutants. 
This is expected as the subtractive clustering does not look at the total amount of data points(mutants) but to the contents of the data points.
The number of clusters compared to the that of the hierarchical clustering is significantly lower.
While we do not know the exact clusters centroids the subtractive model calculated, a low number of clusters indicates close relationships between certain mutants.
\begin{finding}
The mutants represented as characteristics in our data set per project have, according to the subtractive clustering model, close relationships to certain other mutants in the same data set.
\end{finding}
We observe that the minimum lowest minimum cluster accuracy measured over all variants and projects is 50.54\% with and average of 62\%. 
This is higher than the minimum of the hierarchical clustering.
This means that in all project there was not one cluster that was evenly divided.
While the minimum accuracy was higher, the maximum accuracy was lower compared to the maximum accuracy of the hierarchical clustering results.
The highest accuracy measured over all variants and projects is 97.80\% with an average of 83.79\%
This means that in all project there was not one cluster that consisted of only survived or killed mutants.
\begin{finding}
The overall accuracy of the clusters generated by the machine learning model in both variants is lower compared to that of the hierarchical clustering.
\end{finding}
When we compare the accuracy per project between the machine learning model and hierarchical clustering we can observe that the lower accuracy of the machine learning model clusters have a higher impact on the deviation in mutation score.
This is caused by the relatively low number of clusters.
One cluster contains more mutants which results in a bigger weight per cluster.
\begin{finding}
The inaccuracy of clusters has a bigger impact on mutation score when the number of clusters is lower.
\end{finding}

### 2021-07-13 chapters/9_conclusion.tex
number of clusters

\chapter{Conclusion}
\label{ch:conclusion}
Our research started with comparing three mutation testing tools for Java and found that the state of the art was outdated.
We replicated the state of the art research with current data available on the mutation testing tools.
The state of art showed that two out of three mutation testing tools were not actively maintained anymore.
The mutation testing tool PIT scored best on the metrics measured in our replication study, this is also the tool we decided to use for our research.
\newline
We created a list of characteristics based on existing research. 
We then extended this list by identifying characteristic with logic and data available.
We used the identified characteristics to represent the mutants.
\newline
We clustered the mutants with two different clustering techniques; hierarchical clustering and by training a machine learning clustering model.
The hierarchical clustering showed that we can cluster mutants while maintaining effectiveness.
We achieved reductions of up to 75\% of the total amount of mutants without significantly reducing the accuracy.
\newline
The machine learning model showed a significant reduction in effectiveness.
We trained the machine learning model with parameters that were optimal according to existing research. 
We also used a machine learning model, specialized in calculating the number of clusters for fuzzy clustering, to calculate the amount of clusters.
While the minimum clustering accuracy was higher than that of the hierarchical clustering, the maximum cluster accuracy was lower than that of the hierarchical clustering.
The cluster accuracy had a greater impact on the clusters of the machine learning model because there was a relatively small amount of clusters.
Our research proves that it is possible to determine the means of domains as the centre of a cluster within a programming language without narrowing down the scope to specific pieces of code.
It can be applied to all Java projects that can be run with PIT. 
Which is a larger scope than existing mutation clustering research used.

>>>>

\chapter{Conclusion}
\label{ch:conclusion}
Our research started with comparing three mutation testing tools for Java and found that the state of the art was outdated.
We replicated the state of the art research with current data available on the mutation testing tools.
The state of art showed that two out of three mutation testing tools were not actively maintained anymore.
The mutation testing tool PIT scored best on the metrics measured in our replication study, this is also the tool we decided to use for our research.
\newline
We created a list of characteristics based on existing research. 
We then extended this list by identifying characteristic with logic and data available.
We used the identified characteristics to represent the mutants.
\newline
We clustered the mutants with two different clustering techniques; hierarchical clustering and by training a machine learning clustering model.
The hierarchical clustering showed that we can cluster mutants while maintaining effectiveness.
We achieved reductions of up to 75\% of the total amount of mutants without significantly reducing the accuracy.
\newline
The machine learning model showed a significant reduction in effectiveness.
We trained the machine learning model with parameters that were optimal according to existing research. 
We also used a machine learning model, specialized in calculating the number of clusters for fuzzy clustering, to calculate the number of clusters.
While the minimum clustering accuracy was higher than that of the hierarchical clustering, the maximum cluster accuracy was lower than that of the hierarchical clustering.
The cluster accuracy had a greater impact on the clusters of the machine learning model because there was a relatively small number of clusters.
Our research proves that it is possible to determine the means of domains as the centre of a cluster within a programming language without narrowing down the scope to specific pieces of code.
It can be applied to all Java projects that can be run with PIT. 
Which is a larger scope than in existing mutation clustering research.

### 2021-07-13 chapters/rq1/2_experiment_design.tex
fixed our tool to the tool
number of clusters

\subsection{Cluster size}
Agglomerative hierarchical clustering will continue to cluster until there is one cluster left. 
Naturally this cluster will contain all the mutants.
We can cut off the clustering algorithm at any point.
We decide on a cluster amount based on the amount of mutants generated in a full set.
As a baseline we will cut off clustering in steps of 25\%.
The steps of 25\% will give a good global indication in reduction.
If necessary we can decide to cut of the clustering at different points.
The baseline will consist of three experiments where the cluster size is 25\%, 50\% and 75\% of the total amount of mutants respectively.
The amount of mutants inside a cluster will be decided by the clustering algorithm.

\subsection{Validation}
\label{ch:exp1_validation}
The most efficient way to measure test effectiveness with mutation testing is by running all mutants that a tool can possibly generate.
The goal of this experiment is to reduce the amount of mutants executed while maintaining effectiveness.
To reduce the amounts executed we cluster the mutants. 
We can measure this by counting the amount of clusters we generate and compare it to the number of total mutants generated by the selected tool.
\newline
The baseline for our experiment is the mutation score of the set of all mutants generated by our selected mutation testing tool.
\newline
To validate how effective our clustering is we can compare the weighted mutation score(see Chapter \ref{ch:weighted_score}) of our clustered set to the mutation score of the full set.
The closer the weighted score is to that of a full set the more effective our set of characteristics and clustering algorithm proves to be.
In other words we want achieve a mutation score that is as close as possible as to that of a full set.
We select a statistical significance level of $\leq 0.05$.
This is the conventional threshold for declaring statistical significance\cite{Kirk1996PracticalCome}.
\newline
Depending on the effectiveness of our clustering algorithm we may loose accuracy.
This can happen if a cluster contains mutants both types. 
We can measure this accuracy by calculating a percentage of all mutants that have survived against the ones that have been killed in a cluster.
If the majority of the mutants in a cluster is killed then we consider that cluster to represent a killed cluster and the other way around for survived mutants.
We consider the mutants that are not in the majority of the cluster as inaccuracy.
\newline
If the weighted mutation score of our clustered set deviates more than 5\% from the score of a full set we will reject our hypothesis.
We also compare the inaccuracy to the significance level to decide if our inaccuracy is statistical significant.

>>>>

\subsection{Cluster size}
Agglomerative hierarchical clustering will continue to cluster until there is one cluster left. 
Naturally this cluster will contain all the mutants.
We can cut off the clustering algorithm at any point.
We decide on the number of clusters based on the amount of mutants generated in a full set.
As a baseline we will cut off clustering in steps of 25\%.
The steps of 25\% will give a good global indication in reduction.
If necessary we can decide to cut of the clustering at different points.
The baseline will consist of three experiments where the cluster size is 25\%, 50\% and 75\% of the total amount of mutants respectively.
The amount of mutants inside a cluster will be decided by the clustering algorithm.

\subsection{Validation}
\label{ch:exp1_validation}
The most efficient way to measure test effectiveness with mutation testing is by running all mutants that a tool can possibly generate.
The goal of this experiment is to reduce the amount of mutants executed while maintaining effectiveness.
To reduce the amounts executed we cluster the mutants. 
We can measure this by counting the number of clusters we generate and compare it to the number of total mutants generated by the selected tool.
\newline
The baseline for our experiment is the mutation score of the set of all mutants generated by the selected mutation testing tool.
\newline
To validate how effective our clustering is we can compare the weighted mutation score(see Chapter \ref{ch:weighted_score}) of our clustered set to the mutation score of the full set.
The closer the weighted score is to that of a full set the more effective our set of characteristics and clustering algorithm proves to be.
In other words we want achieve a mutation score that is as close as possible as to that of a full set.
We select a statistical significance level of $\leq 0.05$.
This is the conventional threshold for declaring statistical significance\cite{Kirk1996PracticalCome}.
\newline
Depending on the effectiveness of our clustering algorithm we may loose accuracy.
This can happen if a cluster contains mutants both types. 
We can measure this accuracy by calculating a percentage of all mutants that have survived against the ones that have been killed in a cluster.
If the majority of the mutants in a cluster is killed then we consider that cluster to represent a killed cluster and the other way around for survived mutants.
We consider the mutants that are not in the majority of the cluster as inaccuracy.
\newline
If the weighted mutation score of our clustered set deviates more than 5\% from the score of a full set we will reject our hypothesis.
We also compare the inaccuracy to the significance level to decide if our inaccuracy is statistical significant.

### 2021-07-13 chapters/rq2/1_experiment_design.tex
number of clusters

\subsubsection{Neural gas}
The neural gas is a topology-preserving network, and can be treated as an extension to the C-means\cite{Du2010Clustering:Approach}.
A data optimal topological ordering is achieved by using neighborhood ranking within the input space at each training step\cite{Du2010Clustering:Approach}.
To find its neighborhood rank, each neuron compares its distance to the input vector with those of all the other neurons to the input vector\cite{Du2010Clustering:Approach}. 
Unlike the self organizing map, which uses predefined static neighborhood relations, the neural gas network determines a dynamical neighborhood relation as learning proceeds\cite{Du2010Clustering:Approach}.
A big pro of the neural gas network is that it is able to spawn neurons.
This means that we do not need to know the amount of clusters in advance.
Unfortunately the neural gas network is a learning network.
One of the properties of a learning network is that it is not designed to fit new data points\cite{supervisedUnsupervised}.
For this reason we cannot use the neural gas network to predict in what cluster a new mutant should be placed.
Therefore we will not select the neural gas network.

@@

\subsection{Algorithm parameters}
The fuzzy c-means algorithm has to initialize the centroids of the clusters.
Our selected library does this based on a random number.
It also allows us to control this random number.
To make the initialization of the centroid reproducible and deterministic we will use a generated seed.
The seed generated is included in the Github repository\cite{rbasarat-repo}.
\newline
Bezdek et al., also researched cluster validity for the fuzzy c-means model\cite{Bezdek1995OnModel}.
They researched what effect which parameters have on the validity of the clusters.
They specifically analyzed the role of weighting exponent \textit{m}(fuzziness parameter).
They concluded that the best choice for \textit{m} is probably in the interval [1.5, 2.5], whose mean and midpoint is \textit{m}=2.
Wu proposed a new guideline for selecting parameter \textit{m}\cite{Wu2012}.
His point of view was that a large \textit{m} value will make the fuzzy c-means more robust to noise and outliers.
He suggests implementing fuzzy c-means with \textit{m}=[1.5, 4].
When the data set contains noise and outliers, the fuzzifier\textit{m}=4 is recommended in a theoretical upper bound case.
Since our data set does not contain noise and outliers we choose to select the midpoint concluded in the research of Bezdek et al.\cite{Bezdek1995OnModel}.
The value of \textit{m} in this experiment is 2.
\newline
Finding the correct amount of clusters has been an ongoing problem for clustering algorithms\cite{Du2010Clustering:Approach}.
Fortunately there are some techniques available to estimate the cluster amount.
During model selection we discussed the mountain and subtractive models. 
These models are designed to estimate the amount of clusters in a data set.
The subtractive clustering model is specifically designed to approximate the number of clusters for fuzzy system models\cite{Chiu1994FuzzyEstimation}.
It approximates the centroids and amount of the clusters based on the data points in the data set\cite{Chiu1994FuzzyEstimation}.
We will select and implement the subtractive clustering model to estimate the amount of clusters.
We can use this amount for the cluster amount parameter in the fuzzy c-means algorithm.

>>>>

\subsubsection{Neural gas}
The neural gas is a topology-preserving network, and can be treated as an extension to the C-means\cite{Du2010Clustering:Approach}.
A data optimal topological ordering is achieved by using neighborhood ranking within the input space at each training step\cite{Du2010Clustering:Approach}.
To find its neighborhood rank, each neuron compares its distance to the input vector with those of all the other neurons to the input vector\cite{Du2010Clustering:Approach}. 
Unlike the self organizing map, which uses predefined static neighborhood relations, the neural gas network determines a dynamical neighborhood relation as learning proceeds\cite{Du2010Clustering:Approach}.
A big pro of the neural gas network is that it is able to spawn neurons.
This means that we do not need to know the number of clusters in advance.
Unfortunately the neural gas network is a learning network.
One of the properties of a learning network is that it is not designed to fit new data points\cite{supervisedUnsupervised}.
For this reason we cannot use the neural gas network to predict in what cluster a new mutant should be placed.
Therefore we will not select the neural gas network.

@@

\subsection{Algorithm parameters}
The fuzzy c-means algorithm has to initialize the centroids of the clusters.
Our selected library does this based on a random number.
It also allows us to control this random number.
To make the initialization of the centroid reproducible and deterministic we will use a generated seed.
The seed generated is included in the Github repository\cite{rbasarat-repo}.
\newline
Bezdek et al., also researched cluster validity for the fuzzy c-means model\cite{Bezdek1995OnModel}.
They researched what effect which parameters have on the validity of the clusters.
They specifically analyzed the role of weighting exponent \textit{m}(fuzziness parameter).
They concluded that the best choice for \textit{m} is probably in the interval [1.5, 2.5], whose mean and midpoint is \textit{m}=2.
Wu proposed a new guideline for selecting parameter \textit{m}\cite{Wu2012}.
His point of view was that a large \textit{m} value will make the fuzzy c-means more robust to noise and outliers.
He suggests implementing fuzzy c-means with \textit{m}=[1.5, 4].
When the data set contains noise and outliers, the fuzzifier\textit{m}=4 is recommended in a theoretical upper bound case.
Since our data set does not contain noise and outliers we choose to select the midpoint concluded in the research of Bezdek et al.\cite{Bezdek1995OnModel}.
The value of \textit{m} in this experiment is 2.
\newline
Finding the correct amount of clusters has been an ongoing problem for clustering algorithms\cite{Du2010Clustering:Approach}.
Fortunately there are some techniques available to estimate the number of clusters.
During model selection we discussed the mountain and subtractive models. 
These models are designed to estimate the number of clusters in a data set.
The subtractive clustering model is specifically designed to approximate the number of clusters for fuzzy system models\cite{Chiu1994FuzzyEstimation}.
It approximates the centroids and amount of the clusters based on the data points in the data set\cite{Chiu1994FuzzyEstimation}.
We will select and implement the subtractive clustering model to estimate the number of clusters.
We can use this number for the number of clusters parameter in the fuzzy c-means algorithm.
